Bayesian RKHS-based methods in functional regression\support{Support information of the article.}

### 2022-02-19 notes-rkhs.tex
Reorganize and minor corrections

% \documentclass[11pt]{article}
% \usepackage[margin=1.3in]{geometry}
% %\usepackage[compact,small]{titlesec}
% %--------------------------------------------------------
% \usepackage{sectsty,todonotes}
% \sectionfont{\normalsize\bfseries}
% \subsectionfont{\normalsize\sf}
% %----------------------------------------------------
% \usepackage[mathcal]{euscript}
%
% \usepackage{bm}                     %AMS Packages
% \usepackage{amsfonts, framed}
% \usepackage{amssymb}
% \usepackage{amsmath}
% \usepackage{amsthm}
% \usepackage{graphicx}               %Graphics
% %\usepackage{pstricks,pst-plot}
% %\usepackage{wrapfig}
% \usepackage{natbib}                 %Citation style
% \usepackage{colortbl}               %Tables
% \usepackage{booktabs}
% \usepackage{hyperref}
% \setlength{\bibsep}{1pt}
% %\usepackage{hypernat}
%
% % Definiciones
%
% \definecolor{violet}{rgb}{0.7,0,0.6}
% \newcommand{\explainindetail}[1]{\todo[color=violet]{#1}}
% \newcommand{\nat}{\mathbb{N}}
% \newcommand{\eps}{\varepsilon}
% \newcommand{\ind}{\mathbb{I}}
% \newcommand{\pr}{\mathbb{P}}
% \newcommand{\B}{B}
% \newcommand{\A}{\mathcal{A}}\usepackage{amsfonts}
% \usepackage{amssymb}
% \usepackage{amsmath}
% \usepackage{amsthm}
% \newcommand{\E}{\mathcal{E}}
% \newcommand{\V}{\mathbb{V}}
% \newcommand{\F}{\mathfrak{F}}
% \newcommand{\G}{\mathbb{G}}
% \newcommand{\M}{\mathcal{M}}
% \newcommand{\X}{\mathcal{X}}
% \newcommand{\Z}{\mathcal{Z}}
% \newcommand{\T}{\mathcal{T}}
% \newcommand{\Y}{\mathcal{Y}}
% \newcommand{\C}{\mathcal{C}}
% \newcommand{\ds}{\displaystyle}
% \newcommand{\D}{\mathcal{D}}
% \newcommand{\deb}{\stackrel{\mathcal{L}}{\longrightarrow}}
% \newcommand{\Vor}{\mbox{Vor}}
% \newcommand{\ang}{\measuredangle}
% \def\cals_+{{\cals_+}}
% \def\cala{{\mathcal{A}}}
% \def\calb{{\mathcal{B}}}
% \def\calc{{\mathcal{C}}}
% \def\cald{{\mathcal{D}}}
% \def\calf{{\mathcal{F}}}
% \def\calh{{\mathcal{H}}}
% \def\call{{\mathcal{L}}}
% \def\cals{{\mathcal{S}}}
% \def\calv{{\mathcal{V}}}
% \def\calg{{\mathcal{G}}}
% \def\calp{{\mathcal{P}}}
% \def\cale{{\mathcal{E}}}
% \def\calo{{\mathcal{O}}}
% \def\calj{{\mathcal{J}}}
% \def\caln{{\mathcal{N}}}
% \def\calm{{\mathcal{M}}}
%
% \hypersetup{
% colorlinks=true,
% citecolor=violet
% }
%
% \begin{document}

\begin{center}
\large \bf  Preliminaries on RKHS's\rm
\end{center}

RKHS's are spaces of ``regular'' FUNCTIONS (not equivalence classes in $L^2[0,1]$) with a different norm to that inherited from $L^2[0,1]$.

\section*{Definition}

@@

	Let $\calh_0(K):=\{f\in L^2[0,1] \ : \ f(\cdot)=\sum_{i=1}^n a_i K(t_i,\cdot),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\}$, be
	the space of all finite linear combinations of evaluations of $K$.  This space is endowed with the inner product
	$\langle f,g\rangle_K=\sum_{i,j}\alpha_i\beta_j K(t_i,s_j)$,
	where $f(\cdot)=\sum_i\alpha_i K(t_i,\cdot)$ and $g(\cdot)=\sum_j\beta_j K(s_j,\cdot)$.

	Then, the RKHS associated with $K$ is defined as the completion of $\calh_0(K)$. In other words, $\calh(K)$ is made of all functions obtained as pointwise limits of Cauchy sequences in $\calh_0(K)$. The inner product is extended accordingly to the whole space $\calh(K)$.

\section*{Reproducing property}

	These spaces are named after the so-called \textit{reproducing property},
	$\langle f, K(s,\cdot)\rangle_K = f(s)$, for all  $f\in\calh(K), s\in[0,1]$, which is particularly important in the applications. On account of this property it is sometimes said that the RKHS are spaces of ``true functions'', in the sense that the pointwise values $f(s)$ at a given $s$ do matter, in contrast with $L^2[0,1]$ whose elements are in fact equivalence classes of functions.

@@

\section*{Loève's isometry}

	A property of RKHS's especially useful in statistical applications is given by the following \textit{Lo\`eve's isometry}:  let $L^2(\Omega) $ be the Hilbert space of real random variables with finite second moment, endowed with the usual inner product whose associated norm is $\| U\|^2={\mathbb E}(U^2)$. Define
	$$\call_0(X) = \big\{U\in L^2(\Omega) \ : \ U=\sum_{i=1}^n a_i \big(X(t_i)-m(t_i)\big),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\big\},$$
	where $m(t)={\mathbb E}[X(t)]$, and let $\call(X)$ be the completion of $\call_0(X)$ in  $L^2(\Omega)$; hence, in other words, is a closed subspace of $L^2(\Omega)$ defined as the linear span of the centred one-dimensional marginals of the process $X$. It turns out that the transformation $\Psi_X$, from  $\call(X)$ to $\calh(K)$, defined by
	\begin{equation}\label{eq:isometry}
	\Psi_X(U)(s) = {\mathbb E}[ U(X(s)-m(s))] = \langle U, X(s)-m(s) \rangle \in \calh(K), \ \text{ for } U\in \call(X)
	\end{equation}
	is an isometry (sometimes called Lo\`eve's isometry) between $\call(X)$ and $\calh(K)$, that is, $\Psi_X(U)$ is bijective and preserves the inner product. As a consequence, the Hilbert spaces $\call(X)$ and $\calh(K)$ can be identified.

	\textbf{In more informal terms, this isometry is nothing but the completion} of
	$$
	\sum_ia_iK(t_i,\cdot)\longleftrightarrow \sum_ia_i(X(t_i)-m(t_i))
	$$

\section*{Alternative definitions}

@@

\section*{General properties}

  An important thing to note is that in general the trajectories of the process $X$ \textbf{do not belong to the corresponding RKHS}. More precisely, $X(t)\notin \mathcal H(K)$ with probability $1$ if $\mathcal K$ has infinite non-zero eigenvalues. While a rigorous proof of this fact is given in \citet{hajek1962}, an heuristic argument presented in \citet{wahba1990} is as follows. Consider the Karhunen-Loève expansion of $X$, i.e.,
  \[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
  \]

  and the truncated version $X_N(t)$ up to the $N$-th term. On the one hand, $X_N(t)\to X(t)$ in the quadratic mean (for each fixed $t$) by the Karhunen-Loève theorem, but on the other hand we have
  \[
\mathbb E\|X_N(\cdot)\|^2_K = \mathbb E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
  \]

  Thus, functions in the RKHS are in a way ``smoother'' than the trajectories of the process themselves. For convenience, if $x=x(\omega)$ is a realization of $X$ (for a fixed $\omega$), we will denote $\langle x, f\rangle_K \equiv \left[\Psi_X^{-1}(f)\right](\omega)$. Some interesting properties stemming from this interpretation are the following (see \citet[p.~974]{parzen1961b}):

@@

    \item $\langle X, K(t, \cdot)\rangle_K = X(t) - m(t)$

    \item $\mathbb{E} \langle X, h\rangle_K = \langle m, h\rangle_K$

    \item $\operatorname{Cov}\left(\langle X, h \rangle_K , \langle X, g\rangle_K\right) = \langle h, g\rangle_K$

@@

\section*{Convergence properties}

@@

\end{itemize}

\begin{thebibliography}{9}
	\bibitem[\protect\citeauthoryear{Hsing \& Eubank}{2015}]{hsi15}
	Hsing, T., \& Eubank, R. (2015). \textit{Theoretical foundations of functional data analysis, with an introduction to linear operators} (Vol. 997). John Wiley \& Sons.

	\bibitem[\protect\citeauthoryear{Paulsen \& Raghupathi }{2015}]{pau16}
	Paulsen, V. I., \& Raghupathi, M. (2016). \textit{An introduction to the theory of reproducing kernel Hilbert spaces} (Vol. 152). Cambridge University press.
  \bibitem[\protect\citeauthoryear{Parzen}{1961}]{parzen1961b} Parzen, E. (1961). \textit{An Approach to Time Series Analysis}. The Annals of Mathematical Statistics 32(4), 951–989.

  \bibitem[\protect\citeauthoryear{Wahba}{1990}]{wahba1990} Wahba G. (1990). \textit{Spline models for observational data}. CBMS 59, SIAM, Philadelphia.

  \bibitem[\protect\citeauthoryear{Hajek}{1962}]{hajek1962} Hajek J. (1962). \textit{On linear Statistical problems in stochastic processes}. Chekoslovak Math. J., vol. 12, pp.404-443.

\end{thebibliography}

%\end{document}

>>>>

RKHS's are spaces of ``regular'' FUNCTIONS (not equivalence classes in $L^2[0,1]$) with a different norm to that inherited from $L^2[0,1]$. See \sidetextcite{paulsen2016introduction} for a more detailed introduction.

\subsection*{Definition}

@@

	Let $\mathcal H_0(K):=\{f\in L^2[0,1] \ : \ f(\cdot)=\sum_{i=1}^n a_i K(t_i,\cdot),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\}$, be
	the space of all finite linear combinations of evaluations of $K$. This space is endowed with the inner product
	$\langle f,g\rangle_K=\sum_{i,j}\alpha_i\beta_j K(t_i,s_j)$,
	where $f(\cdot)=\sum_i\alpha_i K(t_i,\cdot)$ and $g(\cdot)=\sum_j\beta_j K(s_j,\cdot)$.

	Then, the RKHS associated with $K$ is defined as the completion of $\mathcal H_0(K)$. In other words, $\mathcal H(K)$ is made of all functions obtained as pointwise limits of Cauchy sequences in $\mathcal H_0(K)$. The inner product is extended accordingly to the whole space $\mathcal H(K)$.

\subsection*{Reproducing property}

	These spaces are named after the so-called \textit{reproducing property},
	$\langle f, K(s,\cdot)\rangle_K = f(s)$, for all  $f\in\mathcal H(K), s\in[0,1]$, which is particularly important in the applications. On account of this property it is sometimes said that the RKHS are spaces of ``true functions'', in the sense that the pointwise values $f(s)$ at a given $s$ do matter, in contrast with $L^2[0,1]$ whose elements are in fact equivalence classes of functions.

@@

\subsection*{Loève's isometry}

	A property of RKHS's especially useful in statistical applications is given by the following \textit{Lo\`eve's isometry}:  let $L^2(\Omega) $ be the Hilbert space of real random variables with finite second moment, endowed with the usual inner product whose associated norm is $\| U\|^2={\mathbb E}(U^2)$. Define
	$$\mathcal L_0(X) = \big\{U\in L^2(\Omega) \ : \ U=\sum_{i=1}^n a_i \big(X(t_i)-m(t_i)\big),\ a_i\in{\mathbb R},\ t_i\in[0,1],\ n\in{\mathbb N}\big\},$$
	where $m(t)={\mathbb E}[X(t)]$, and let $\mathcal L(X)$ be the completion of $\mathcal L_0(X)$ in  $L^2(\Omega)$; hence, in other words, is a closed subspace of $L^2(\Omega)$ defined as the linear span of the centred one-dimensional marginals of the process $X$. It turns out that the transformation $\Psi_X$, from  $\mathcal L(X)$ to $\mathcal H(K)$, defined by
	\begin{equation}\label{eq:isometry}
	\Psi_X(U)(s) = {\mathbb E}[ U(X(s)-m(s))] = \langle U, X(s)-m(s) \rangle \in \mathcal H(K), \ \text{ for } U\in \mathcal L(X)
	\end{equation}
	is an isometry (sometimes called Lo\`eve's isometry) between $\mathcal L(X)$ and $\mathcal H(K)$, that is, $\Psi_X(U)$ is bijective and preserves the inner product. As a consequence, the Hilbert spaces $\mathcal L(X)$ and $\mathcal H(K)$ can be identified.

	In more informal terms, this isometry is nothing but the completion of
	$$
	\sum_ia_iK(t_i,\cdot)\longleftrightarrow \sum_ia_i(X(t_i)-m(t_i))
	$$

\subsection*{Alternative definitions}

@@

\subsection*{General properties}

An important thing to note is that in general the trajectories of the process \(X\) \textbf{do not belong to the corresponding RKHS}. More precisely, \(X(t)\notin \mathcal H_k\) with probability \(1\) if \(\mathcal K\) has infinite non-zero eigenvalues. While a rigorous proof of this fact is given in \sidetextcite{hajek1962linear}, an heuristic argument presented in \sidetextcite{wahba1990spline} is as follows. Consider the Karhunen-Loève expansion of \(X\), i.e.,
\[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
\]
and the truncated version \(X_N(t)\) up to the \(N\)-th term. On the one hand, \(X_N(t)\to X(t)\) in the quadratic mean (for each fixed \(t\)) by the Karhunen-Loève theorem, but on the other hand we have
\[
  \mathds E\|X_N(\cdot)\|^2_k = \mathds E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
\]

Thus, functions in the RKHS are in a way \textit{smoother} than the trajectories of the process themselves. For convenience, if \(x=x(\omega)\) is a realization of \(X\) (for a fixed \(\omega\)), we will denote \(\langle x, f\rangle_k \equiv \left[\Psi_X^{-1}(f)\right](\omega)\). Some interesting properties stemming from this interpretation are the following (see \sidetextcite[][p.~974]{parzen1961approach}):

@@

  \item \(\langle X, k(\cdot, t)\rangle_k = X(t) - m(t)\).

  \item \(\mathds{E} \langle X, f\rangle_k = \langle m, f\rangle_k\).

  \item \(\cov \left(\langle X, f \rangle_k , \langle X, g\rangle_k\right) = \langle f, g\rangle_k\).

@@

\subsection*{Convergence properties}

@@

### 2022-02-19 notes.tex
Reorganize and minor corrections

\section{TODO}

\item Why does it not make sense to regress \(Y\) on a finite combination of projections of \(X\) whenever we consider the \(L^2\) model?

\textbf{ANSWER}: Continuous linear functionals are only obtained by integrating a coefficient function against $x(t)$ (Riesz-Fisher), and it is not possible to select a $\beta(t)$ that produces finite linear combinations of the process after integrating.

\item \(\alpha(t)=E [X(t)U]\) is the correlation/covariance of $X$ and $U$ (Berrendero et al. 2019, 2020), if we assume that $E[X(t) \varepsilon]=0$. Moreover, are we assuming $E[(X(t) - m(t))U)]=0$?

\item Since the RMSE is measured on the same scale, with the same units as y, one can expect 68\% of the y values to be within 1 RMSE — given the data is normally distributed.

@@

\marginnote{We assume the functional regressor \(X=X(t)\) to be an \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), a vanishing mean function and a continuous covariance function \(k=k(s, t)\).}

We consider the RKHS-based functional linear regression model given by\sidenote[][*7]{This formulation comes from considering the (centered) model \(Y=\Psi_X^{-1}(\alpha) + \eps\), where \(\eps\) is an independent normal error term with \(\E\eps=0\) and \(\var(\eps)=\sigma^2\).}
\begin{equation}
  \label{eq:model_lin}
  Y_i \mid \theta, X_i \stackrel{\text{i.i.d.}}{\sim} \mathcal N(\Psi_{X_i}^{-1}(\alpha), \sigma^2),
\end{equation}
where \(\Psi_X\) is \textit{Loeve's isometry} and \(\alpha \in \mathcal H_{k}\) is the regression coefficient function. The main idea is to restrict \(\alpha\) to the linear span of a finite subset of evaluations of the kernel \(\{k(\cdot, t_j)\}_j\), since the whole set of such evaluations is dense in \(\mathcal H_k\). Specifically, we only consider functions of the form
\begin{equation}
  \label{eq:alpha}
  \alpha(t) = \sum_{j=1}^p \beta_j k(t, t_j),
\end{equation}
where \(p\) is a fixed hyperparameter. Under these assumptions, the parametric space is the set \(\Theta \subseteq \R^p \times [0, 1]^p \times \R^+\), with elements
\[
  \theta = (\beta_1, \dots, \beta_p, t_1,\dots, t_p, \sigma^2)' \equiv (\beta, \tau, \sigma^2)'.
\]

Following a Bayesian approach, we can set a prior distribution on the parameters, such as:
\begin{align*}
  f(\sigma^2)              & \propto 1/\sigma^2,                                                     \\
  \tau                     & \sim \mathscr U([0, 1]^p),                                              \\
  \beta\mid \tau, \sigma^2 & \sim \mathcal N(b_0, g\sigma^2(\mathcal X_\tau' \mathcal X_\tau)^{-1}),
\end{align*}
where\sidenote[][*1]{This is because of the correspondence between \(k(\cdot, t_j)\) and \(X(t_j)\), which becomes relevant when plugging the approximation~\eqref{eq:alpha} into the model~\eqref{eq:model_lin}.}
\(\mathcal X_\tau(i, j)=X_i(t_j)\). The prior on \(\beta\) is known as Zellner's \(g\)-prior owing to the main hyperparameter \(g\in \R^+\), which controls the confidence we have in the initial estimate \(b_0 \in \R^p\) for the true value of the coefficients. The choice of a non-informative improper prior for \(\sigma^2\) (known as Jeffreys prior) is of course subject to change. All in all, the prior distribution for the parameters becomes
\begin{equation}
  \label{eq:prior}
  \pi(\theta) = \pi(\beta, \tau, \sigma^2) = \pi(\beta \mid \tau, \sigma^2)\pi(\tau \mid \sigma^2)\pi(\sigma^2).
\end{equation}

@@

@@

An important thing to note is that in general the trajectories of the process \(X\) \textbf{do not belong to the corresponding RKHS}. More precisely, \(X(t)\notin \mathcal H_k\) with probability \(1\) if \(\mathcal K\) has infinite non-zero eigenvalues. While a rigorous proof of this fact is given in \sidetextcite{hajek1962linear}, an heuristic argument presented in \sidetextcite{wahba1990spline} is as follows. Consider the Karhunen-Loève expansion of \(X\), i.e.,
\[
  X(t) = \sum_{j=1}^\infty \zeta_j \phi_j(t),
\]
and the truncated version \(X_N(t)\) up to the \(N\)-th term. On the one hand, \(X_N(t)\to X(t)\) in the quadratic mean (for each fixed \(t\)) by the Karhunen-Loève theorem, but on the other hand we have
\[
  \mathds E\|X_N(\cdot)\|^2_k = \mathds E \sum_{j=1}^N \frac{\zeta_j^2}{\lambda_j} = N \to \infty \quad (N\to\infty).
\]

Thus, functions in the RKHS are in a way \textit{smoother} than the trajectories of the process themselves. For convenience, if \(x=x(\omega)\) is a realization of \(X\) (for a fixed \(\omega\)), we will denote \(\langle x, f\rangle_k \equiv \left[\Psi_X^{-1}(f)\right](\omega)\). Some interesting properties stemming from this interpretation are the following (see \sidetextcite[][p.~974]{parzen1961approach}):

  \item \(\langle X, k(\cdot, t)\rangle_k = X(t) - m(t)\).

  \item \(\mathds{E} \langle X, f\rangle_k = \langle m, f\rangle_k\).

  \item \(\cov \left(\langle X, f \rangle_k , \langle X, g\rangle_k\right) = \langle f, g\rangle_k\).

%\include{notes-rkhs.tex}

>>>>

\section{Functional Data Analysis}

See \sidetextcite{hsing2015theoretical} for a complete discussion on the subject.

@@

\marginnote{We assume the functional regressor \(X=X(t)\) to be an \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), a vanishing mean function and a continuous covariance function \(K=K(t, s)\).}

We consider the RKHS-based functional linear regression model given by\sidenote[][*7]{This formulation comes from considering the (centered) model \(Y=\Psi_X^{-1}(\alpha) + \eps\), where \(\eps\) is an independent normal error term with \(\E\eps=0\) and \(\var(\eps)=\sigma^2\).}
\begin{equation}
  \label{eq:model_lin}
  Y_i \mid \theta, X_i \stackrel{\text{i.i.d.}}{\sim} \mathcal N(\Psi_{X_i}^{-1}(\alpha), \sigma^2),
\end{equation}
where \(\Psi_X\) is \textit{Loeve's isometry} and \(\alpha \in \mathcal H(K)\) is the regression coefficient function. The main idea is to restrict \(\alpha\) to the linear span of a finite subset of evaluations of the kernel \(\{K(\cdot, t_j)\}_j\), since the whole set of such evaluations is dense in \(\mathcal H(K)\). Specifically, we only consider functions of the form
\begin{equation}
  \label{eq:alpha}
  \alpha(t) = \sum_{j=1}^p \beta_j K(t, t_j),
\end{equation}
where \(p\) is a fixed hyperparameter. Under these assumptions, the parametric space is the set \(\Theta \subseteq \R^p \times [0, 1]^p \times \R^+\), with elements
\[
  \theta = (\beta_1, \dots, \beta_p, t_1,\dots, t_p, \sigma^2)' \equiv (\beta, \tau, \sigma^2)'.
\]

Following a Bayesian approach, we can set a prior distribution on the parameters, such as:
\begin{align*}
  f(\sigma^2)              & \propto 1/\sigma^2,                                                     \\
  \tau                     & \sim \mathscr U([0, 1]^p),                                              \\
  \beta\mid \tau, \sigma^2 & \sim \mathcal N(b_0, g\sigma^2(\mathcal X_\tau' \mathcal X_\tau)^{-1}),
\end{align*}
where\sidenote[][*1]{This is because of the correspondence between \(K(\cdot, t_j)\) and \(X(t_j)\), which becomes relevant when plugging the approximation~\eqref{eq:alpha} into the model~\eqref{eq:model_lin}.}
\(\mathcal X_\tau(i, j)=X_i(t_j)\). The prior on \(\beta\) is known as Zellner's \(g\)-prior owing to the main hyperparameter \(g\in \R^+\), which controls the confidence we have in the initial estimate \(b_0 \in \R^p\) for the true value of the coefficients. The choice of a non-informative improper prior for \(\sigma^2\) (known as Jeffreys prior) is of course subject to change. All in all, the prior distribution for the parameters becomes
\begin{equation}
  \label{eq:prior}
  \pi(\theta) = \pi(\beta, \tau, \sigma^2) = \pi(\beta \mid \tau, \sigma^2)\pi(\tau \mid \sigma^2)\pi(\sigma^2).
\end{equation}

@@

\input{notes-rkhs.tex}

@@

### 2022-02-19 paper/paper.tex
Reorganize and minor corrections

\section{Introduction}

@@

The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces. In what follows, suppose \(X=X(t)\) is an \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its mean function \(m(t)=\mathbb E[X(t)]\) and the covariance function \(K(s, t)= \mathbb E[(X(t) - m(t)(X(s) - m(s)))]\) are both continuous. To construct the RKHS associated with the covariance function, we start by defining the functional space \(H_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
\[
H_0(K) = \left\{ f \in L^2[0,1]: f(\cdot) = \sum_{i=1}^p a_i K(\cdot, t_i), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
\]
This space is endowed with the inner product
\[
\langle f, g\rangle_K = \sum_{i, j} a_i b_j K(s_i, t_j),
\]
given that \(f(\cdot)=\sum_i a_i K(\cdot, s_i) \) and \(g(\cdot)=\sum_j b_j K(\cdot, t_j)\). Lastly, we define \(H(K)\) to be the completion of \(H_0(K)\) under the norm induced by the scalar product defined above.

As we can see, the covariance function plays a crucial role in characterizing the RKHS. An operator closely related to this covariance function is the so-called covariance operator, i.e.:
\[
\mathcal Kf(\cdot) = \int_0^1 K(\cdot, t)f(t)\, dt.
\]
This operator is especially relevant due to the \textit{Karhunen-Loève} theorem \incomment{cita}, which asserts that \(X\) can be decomposed as
\[
X(t) = \sum_{j=1}^\infty \xi_j \phi_j(t),
\]
where \(\{\phi_j\}\) are the eigenfunctions of the covariance operator \(\mathcal K\) and \(\xi_j = \int X\phi_j\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_j\), the eigenvalue associated with \(\phi_j\):

@@

\begin{enumerate}
  \item Loeve's isometry
  \item Sum with lambda j
  \item Square root of K

\end{enumerate}3

\subsection{Overview of Markov chain Monte Carlo methods}

@@

Note that starting from a probability distribution \(\mathbb{P}_0\) on \(H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap H_0(K))\) for all Borel sets \(B\). The following result proves that any such extension is indeed unique.

@@

@@

@@

\bibliographystyle{ba}
\bibliography{sample}

>>>>

\section{Introduction}\label{sec:intro}

The problem of inferring a scalar response \(Y\) from a functional covariate \(X(t)\)

\subsubsection{Organization of the paper}

In Section~\ref{sec:intro} we review...

@@

The methodology proposed in this work relies heavily on the use of RKHS's, so before diving into it we will briefly describe the main characteristics of these spaces. In what follows, suppose \(X=X(t)\) is an \(L^2\)-process with trajectories in \(L^2[0,1]\), and further assume that its mean function \(m(t)=\mathbb E[X(t)]\) and its covariance function \(K(t, s)= \mathbb E[(X(t) - m(t))(X(s) - m(s))]\) are both continuous. To construct the RKHS associated with the covariance function, we start by defining the functional space \(\mathcal H_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
\[
\mathcal H_0(K) = \left\{ f \in L^2[0,1]: f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
\]
This space is endowed with the inner product
\[
\langle f, g\rangle_K = \sum_{i, j} a_i b_j K(t_i, s_j),
\]
given that \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\). Lastly, we define \(\mathcal H(K)\) to be the completion of \(\mathcal H_0(K)\) under the norm induced by the scalar product defined above.

\begin{comment}
  \(K\) is positive and symmetric. We will refer to it as a \textit{kernel}.
\end{comment}

\begin{comment}
  Reproducing property? Convergence entails uniform convergence?
\end{comment}

As we can see, the covariance function plays a crucial role in characterizing the RKHS. A compact operator closely related to this covariance function is the so-called covariance operator, i.e.:
\[
\mathcal Kf(\cdot) = \int_0^1 K(s, \cdot)f(s)\, ds, \quad f \in L^2[0, 1].
\]
This operator is especially relevant due to the \textit{Karhunen-Loève} theorem \incomment{cita}, which asserts that \(X\) can be decomposed as
\begin{equation}\label{eq:karhunen-loeve}
X(t) = \sum_{j=1}^\infty \xi_j \phi_j(t),
\end{equation}
where \(\phi_j\) are the (orthonormal) eigenfunctions of the covariance operator \(\mathcal K\) and \(\xi_j = \int X\phi_j\) are uncorrelated zero-mean random variables with variance equal to \(\lambda_j\), the eigenvalue associated with \(\phi_j\). This expansion is the main justification of many FDA techniques, such as functional principal component analysis \incomment{cita}.

@@

\begin{enumerate}
  \item Via \textit{Loeve's isometry}, one can establish a congruence between \(\mathcal H(K)\) and the linear span of the centered process \(X - m\) in the space of all random variables with finite second moment, \(L^2(\Omega)\) \incomment{cita}. This isometry is essentially the completion of the correspondence
  \begin{equation}\label{eq:loeve}
  \sum_{i=1}^p a_i (X(t_i) - m(t_i)) \longleftrightarrow \sum_{i=1}^p a_i K(t_i, \cdot).
\end{equation}
  \item The space \(\mathcal H(K)\) can also be defined as the image of the square root of the covariance operator, i.e.:
  \begin{equation}\label{eq:rkhs-square-root}
  \mathcal H(K) = \mathcal K^{1/2}(L^2[0, 1]).
\end{equation}

  \item Lastly, starting from the above definition and combining expression~\eqref{eq:karhunen-loeve} and the representation of \(K\) given by \textit{Mercer's theorem} \incomment{cita}, we can also express
  \begin{equation}\label{rkhs-sum-lambda}
    \mathcal H(K) = \left\{f \in L^2[0, 1]: \sum_{j=1}^\infty \frac{\langle f, \phi_j\rangle}{\lambda_j} < \infty \right\},
  \end{equation}
  with the corresponding inner product
  \[
  \langle f, g\rangle_K = \sum_{j=1}^\infty \frac{\langle f, \phi_j\rangle\langle g, \phi_j\rangle}{\lambda_j}.
  \]
Note that, since the spectral theorem for compact operators tells us that \(\lambda_j \to 0\) \incomment{cita}, this definition highlights the fact that functions in \(\mathcal H(K)\) are smooth, in the sense that their components in an orthonormal basis need to vanish quickly \incomment{and?}.
\end{enumerate}

\begin{comment}
  Trajectories not in RKHS
\end{comment}

\begin{comment}
  Small RKHS norm implies small variations and hence a higher degree of smoothness. Specifically, \(\|f\|_K\) controls how fast the function varies over the input space with respect to the geometry defined by the kernel:
  \[
  |f(t) - f(s)| = |\langle f, K_t - K_s\rangle_K| \leq \|f\|_K \|K_t - K_s\|_K.
  \]
\end{comment}

@@

The general idea explored in this work revolves around imposing a prior distribution on the functional space \(\mathcal H_0(K)\), which we know is dense in \(\mathcal H(K)\). Specifically, we consider a functional parameter of the form
\begin{equation}\label{eq:parameter-rkhs}
  \beta(t) = \sum_{i=1}^p \beta_i K(t_i, \cdot),
\end{equation}
and regard \(p, \beta_i\) and \(t_i\) as free parameters in the parameter space \(\Theta\).

Note that starting from a probability distribution \(\mathbb{P}_0\) on \(\mathcal H_0(K)\) we can obtain a probability distribution \(\mathbb{P}\) on \(\mathcal H(K)\) simply by defining \(\mathbb{P}(B) = \mathbb{P}_0(B\cap \mathcal H_0(K))\) for all Borel sets \(B\). The following result proves that any such extension is indeed unique.

@@

\begin{comment}
  Note that in an RKHS the evaluation functional is continuous, so a small perturbation \(f\mapsto f + \delta f\) results in a small change in \(f(x)\) for all \(x\). It makes sense to choose the parameter in \(\mathcal H(K)\). Also smooth functions and linear: the model is \textit{linear} because of the correspondence between the RKHS and the linear span of the process (Loeve's isometry).
\end{comment}

\begin{comment}
  Overview of MCMC methods.
\end{comment}

@@

Berrendero et al. (2020).

@@

\bibliographystyle{ba}
\bibliography{bibliography}
